### 2008-03-18 trunk/Papers/dl-gre-08/discussion.tex
example outputs

Now, we will compare the descriptions generated by our algorithm to
those humans produce. For this purpose, we use a corpus of
human-generated referring expressions collected and made available by
Jette Viethen and Robert
Dale\footnote{http://www.ics.mq.edu.au/~jviethen/drawers}.  The data
was collected in an experiment where participants were asked to
describe one of 16 filing cabinet drawers. The drawers had different
colors and were arranged in a four-by-four grid, as shown in Figure
\ref{fig:drawers}. The human-generated descriptions use four
non-relational (the drawer's \textsf{color}, its \textsf{column} and
\textsf{row} number, and whether it is in a \textsf{corner}) and five
relational properties (\textsf{above, below, next to, left of, right
of}). Of the 118 referring expressions, only 15 use relational properties.
\cite{viethen06:_algor_for_gener_refer_expres}
describe the data in more detail and present results of evaluating the
Full Brevity algorithm by \cite{}, the Incremental Algorithm by
\cite{} and the relational algorithm by \cite{} on this corpus.

@@

As described above \todo{needs to be described}, Dale and Reiter's
Incremental Algorithm is a special case of our algorithm, where no
roles are used and a specific order is imposed on the predicates. We
replicated Viethen and Dale's experiment, which tested the Incremental
Algorithm on the drawer domain with all possible preference ordering
given the four predicates
\textsf{color, column, row}, and \textsf{corner}
\cite{viethen06:_algor_for_gener_refer_expres}.  Unsurprisingly, we
get the same results as they did: 98 of the 103 non-relational
descriptions produced by humans can be generated using one of the 24
possible preference orderings.

\newcite{viethen06:_algor_for_gener_refer_expres} also tested Dale and
Haddock's \shortcite{dale91:_gener_refer_expres_invol_relat}
Relational Algorithm on the drawer domain. Surprisingly, it could not
generate \textit{any} of the descriptions produced by humans. To
evaluate our algorithm's use of relations, we tested it on a
representation of the domain using the predicates \textsf{color} and
\textsf{corner} and the roles \textsf{above, below, next to, left of},
and \textsf{right of}. We did not include the \textsf{column} and
\textsf{row} predicates because every drawer can be uniquely
identified through those predicates. Since our algorithm has to
consider predicates before relations, this would mean that no relation
ever gets used. Like for the Incremental Algorithm, we tested all
possible orderings of predicates and roles (with the restriction that
all predicates come before all roles). Our algorithm generates 10 of
the 15 relational descriptions. Of the five that it could not
generate, two mentioned column information, which we excluded from our
domain, and three (which interestingly were all produced by the same
person) are of the format \textit{the blue/pink/yellow/orange drawer
above/below/next to/right of/left of the two blue/pink/yellow/orange
drawers}. This would correspond to the following output from our
algorithm: $C_1 \& \exists R . (C_2 \& \exists R . C_2)$ (where $C_i$
are color predicates and $R$ a role). While our algorithm does not
generate descriptions that contain the same relation on two levels of
embedding for these drawers, it does generate some other, simpler
descriptions for them; for example, $\textsf{orange} \& \exists
\textsf{left\_of} . (\textsf{pink}
\& \textsf{corner})$ which could be rendered as
\textit{the orange drawer left of the pink one in the corner}, or
$\textsf{orange} \& \exists \textsf{below} . \textsf{yellow}$,
which could be rendered as \textit{the orange drawer below the yellow
drawer}.

@@

@@

\begin{itemize}
\item there's a risk to generate a concept that can't be realized,
  especially for the \alc\ algorithm
\item everybody except perhaps SPUD has this problem
\item problem is worse in our case because it is harder to control the
  order in which atoms and relations are explored
\item would be interesting for future work to think about how
  search in the bisim algorithms can be controlled by the realizer
\end{itemize}

>>>>

To compare the descriptions generated by our algorithm to
those humans produce, we use a corpus of
human-generated referring expressions collected and made available by
Jette Viethen and Robert
Dale\footnote{http://www.ics.mq.edu.au/~jviethen/drawers}.  The data
was collected in an experiment where participants were asked to
describe one of 16 filing cabinet drawers. The drawers had different
colors and were arranged in a four-by-four grid, as shown in Figure
\ref{fig:drawers}. The human-generated descriptions use four
non-relational properties (the drawer's \textsf{color}, its \textsf{column} and
\textsf{row} number, and whether it is in a \textsf{corner}) and five
relational properties (\textsf{above, below, next to, left of, right
of}). Of the 118 referring expressions, only 15 use relational properties.
Viethen and Dale \shortcite{viethen06:_algor_for_gener_refer_expres}
describe the data in more detail and present results of evaluating the
Full Brevity algorithm by \cite{}, the Incremental Algorithm by
\cite{} and the relational algorithm by \cite{} on this corpus.

@@

As described above \todo{needs to be described}, Dale and Reiter's
Incremental Algorithm is a special case of our algorithm, where no
roles are used and a specific order is imposed on the predicates. We
replicated Viethen and Dale's experiment, which tested the Incremental
Algorithm on the drawer domain with all possible preference ordering
given the four predicates
\textsf{color, column, row}, and \textsf{corner}
\cite{viethen06:_algor_for_gener_refer_expres}.  Unsurprisingly, we
get the same results as they did: 98 of the 103 non-relational
descriptions produced by humans can be generated using one of the 24
possible preference orderings. Unfortunately, there isn't one ordering
that covers all those 98 descriptions - a set of four different
orderings is necessary.

\newcite{viethen06:_algor_for_gener_refer_expres} also tested Dale and
Haddock's \shortcite{dale91:_gener_refer_expres_invol_relat}
Relational Algorithm on the drawer domain. Surprisingly, it could not
generate \textit{any} of the descriptions produced by humans. To
evaluate our algorithm's use of relations, we tested it on a
representation of the domain using the predicates \textsf{color} and
\textsf{corner} and the roles \textsf{above, below, next to, left of},
and \textsf{right of}. We did not include the \textsf{column} and
\textsf{row} predicates because every drawer can be uniquely
identified through those predicates. Since our algorithm has to
consider predicates before relations, this would mean that no relation
ever gets used. Like for the Incremental Algorithm, we tested all
possible orderings of predicates and roles (with the restriction that
all predicates come before all roles). For 10 of the 15 human-produced
relational descriptions there is an ordering such that using this
ordering our algorithm can generate this description. Figure
\ref{fig:example_outputs} shows the output of our \el\ algorithm for
one preference ordering of predicates. To cover 10 of the 15
human-produced descriptions three different orderings are
necessary. Of the five human-produced escriptions that the \el\
algorithm can not generate, two mention column information, which we
excluded from our domain, and three (which interestingly were all
produced by the same person) are of the format \textit{the
blue/pink/yellow/orange drawer above/below/next to/right of/left of
the two blue/pink/yellow/orange drawers}. This would correspond to the
following output from our algorithm: $C_1 \& \exists R . (C_2 \&
\exists R . C_2)$ (where $C_i$ are color predicates and $R$ a
role). While our algorithm does not generate descriptions that contain
the same relation on two levels of embedding for these drawers, it
does generate some other, simpler descriptions for them; for example,
$\textsf{orange} \& \exists
\textsf{left\_of} . (\textsf{pink}
\& \textsf{corner})$ which could be rendered as
\textit{the orange drawer left of the pink one in the corner}, or
$\textsf{orange} \& \exists \textsf{below} . \textsf{yellow}$,
which could be rendered as \textit{the orange drawer below the yellow
drawer}.

@@

\begin{figure*}

\begin{small}
\begin{tabular}{|lp{0.5\textwidth}p{0.4\textwidth}|}\hline
drawer & human-produce description & output of the \el\ algorithm\\\hline
2 & \textit{the orange drawer above the blue drawer} & $\exists \textsf{above}.(\textsf{blue}) \& \textsf{orange}$\\
4 & \textit{the yellow drawer on the top of the pink one} & $\exists \textsf{above}.(\textsf{pink}) \& \textsf{yellow}$\\
5 & \textit{the pink drawer in the fourth column below the yellow one} & $\exists \textsf{above}.(\textsf{orange}) \& \textsf{pink}$\\
6 & \textit{the yellow drawer on top of the yellow drawer} & $\exists \textsf{above}.(\textsf{yellow}) \& \textsf{yellow}$\\
6 & \textit{the yellow drawer that's above another yellow drawer} & \\
6 & \textit{the drawer after the two blue ones in horizontal sequence} & \\
7 & \textit{the blue drawer below the orange drawer in the second column} & $\exists \textsf{next}.(\exists \textsf{above}.(\textsf{orange}) \& \textsf{blue}) \& \exists \textsf{above}.(\textsf{blue}) \& \textsf{blue}$\\
7 & \textit{the blue drawer below the orange one} & \\
10 & \textit{the blue drawer above the pink drawer} & $\exists \textsf{above}.(\textsf{pink}) \& \textsf{blue}$\\
10 & \textit{the blue drawer above the pink drawer} & \\
11 & \textit{the yellow drawer next to the orange drawer} & $\exists \textsf{above}.(\textsf{orange}) \& \textsf{yellow}$\\
11 & \textit{the yellow one next to the orange one} & \\
12 & \textit{the orange drawer below the pink drawer} & $\exists \textsf{above}.(\textsf{pink} \& \textsf{corner}) \& \textsf{orange}$\\
14 & \textit{the orange drawer below the two yellow drawers} & $\exists \textsf{next}.(\textsf{pink} \& \textsf{corner}) \& \exists \textsf{next}.(\textsf{pink}) \& \textsf{orange}$\\
14 & \textit{the orange drawer below the two yellow drawers} & \\\hline
\end{tabular}
\end{small}

\caption{Human produced descriptions and output of our \el\  algorithm using the ordering \textsf{corner}, \textsf{color}, \textsf{above, next, below, right, left}}
\label{fig:example_outputs}
\end{figure*}

% ordering no. 235
%List(corner, green, pink, orange, blue)
%List(above, next, below, right, left)

@@

%\begin{itemize}
%\item there's a risk to generate a concept that can't be realized,
%  especially for the \alc\ algorithm
%\item everybody except perhaps SPUD has this problem
%\item problem is worse in our case because it is harder to control the
%  order in which atoms and relations are explored
%\item would be interesting for future work to think about how
%  search in the bisim algorithms can be controlled by the realizer
%\end{itemize}

Like other algorithms that do content determination without taking
into account linguistic contraints, we have no guarantee that the
descriptions generated by our algorithm can actually be realized. For
example, none of the descriptions out algorithms generate in the
drawer domain contain a property which would commonly be realized as a
noun (the property \textsf{drawer} is not included because it applies
to all entities in this domain). If we fed this output directly into a
realizer, the realizer would thus be faced with the task of having to
produce a noun phrase without being given anything that could serve as
the head noun. Of course, this particular problem could easily be
dealt with through a post-processing step that adds a type property to
the description. But at least for the \alc\ algorithm it is not clear
that all non-realizable descriptions could be fixed through such
simple post-processing solutions.

Partly in response to this problem, Horacek \shortcite{} proposes to
interleave content determination with realization and build the
linguistic representation of a natural language description in
parallel with building the representation of its semantic content.

SPUD \cite{} is another system that avoids this problem by
interleaving content determination with realization. SPUD goes one
step further and instead of generating referring expressions in
isolation, deals with them in the context of the sentence they appear
in taking into account semantic information contributed by other parts
of that sentence. This allows them to generate utterances like
\textit{``remove the rabbit from the hat''} which can be used in contexts
containing several rabbits and several hats as long as there is only
one rabbit-hat pair where the rabbit is in the hat. While our
algorithm does generate referring expressions for all entities in the
context at once (and thereby avoids the problem of infinite regress
\todo{discussed earlier}), it is currently not designed to take into
account additional information provided by the context. When used to
generate the expressions referring to the relevant rabbit and hat in
the situation described above, our algorithms would produce
\textit{``the rabbit in the hat''} and \textit{``the hat with the
rabbit''}, respectively, yielding \textit{``remove the rabbit in the
hat from the hat with the rabbit''}.

To improve the interface between our algorithm and a realization
component, we therefore need to investigate ways of how linguistic
contraints on realization can be used to guide our content
determination algorithms and how it can be integrated into the process
of generating a whole utterance in such a way that relevant
information contributed by other parts of the utterance can be taken
into account.
